---
title: "Lab 3 Report"
author: "Kelsey Leach"
date: "`r format(Sys.Date())`"
output: github_document
---

Instructions for this lab report can be found on [Your assignment section of Lab Assignment 3 Introduction to ggplot2](https://biol275-msum.github.io/introduction-to-ggplot2.html#your-assignment) on the lab website.

## Fireflies

> A. Insert an R code chunk and create a graph depicting the frequency distribution of the 35 mass measurements. It should have legible text and appropriate axis labels.

```{r}
#This section brings in the firefly data, counts the number of fireflies for each mass, and creates a histogram with the data.
library(tidyverse)
firefly_data <- read_csv("https://whitlockschluter.zoology.ubc.ca/wp-content/data/chapter02/chap02q19FireflySpermatophoreMass.csv")

distinct(firefly_data, spermatophoreMass)
count(firefly_data, spermatophoreMass)

ggplot(data = firefly_data) + 
  geom_histogram(mapping = aes(x = spermatophoreMass), binwidth = 0.01, 
                 boundary = 0, closed = "left",
                 fill = "#C5351B", color = "black") +
  labs(x = "Spermatophore Mass (mg)", y = "Frequency (number of individuals)") +
  scale_y_continuous(breaks = seq(0, 12, 2), limits = c(0, 12),
                     expand = expansion(mult=0)) +
  scale_x_continuous(breaks = seq(0, 0.2, 0.02)) +
  theme_classic() + 
  theme(
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black", size = rel(1))
  )


```




> B. What type of graph did you choose in part (A)? Why?

I chose to use a histogram in part A because a histogram can be used to display the distribution
of a single numerical variable, in this case, spermatophore mass.

> C. Describe the shape of the frequency distribution. What are its main features?

The frequency appears bimodal with peaks at approximately 0.4 and 0.6. There is an outlier at 0.17. 

> D. What term would be used to describe the largest measurement in the frequency distribution? You can refer to the lecture slides, your notes, or the textbook.

The measurement with the highest frequency would be the mode. 

## Bird orders

I read the data using the code provided by the instructor. The dataset is from the auk package and was originally named ebird_taxonomy. I converted it to a tibble, removed non-species taxa, and named the new dataset birds.

```{r}
#This section brings in the bird data and removes the non-species taxa. 
library(auk)                          
birds <- ebird_taxonomy %>%           
  as_tibble() %>%                     
  filter(category == "species")       
```

> E. How many bird species are in the new birds dataset? How did you find out?

There are 10,721 bird species in the dataset. I found this by looking at the number of rows in the birds dataset after it was printed. 

> H. How many orders are there? You can get this right by counting on the graph, but you will only get full points if you use some code to figure it out.

There are 41 bird orders. This can be determined by looking at the number of rows returned by the
distinct function. 
```{r}
#This section prints all of the bird data and then creates a list of all of the orders present in the data.
birds
distinct(birds, order)
```


> F. Insert an R code chunk and create a graph depicting the distribution of orders in the birds dataset. Sort the orders with the most frequent on the left. It should have legible text and appropriate axis labels.

```{r fig.height=5, fig.width=9}
#This section creates a bar graph with the bird order data. 
ggplot(data = birds) +
  geom_bar(mapping = aes(x=fct_infreq(order)), fill = "#C5351B", 
           width = 0.8) +
  labs(x = "Order", y = "Frequency (number of species)") +
  scale_y_continuous(breaks = seq(0, 6500, 1000), limits = c(0, 6500), 
                     expand = expansion(mult = 0)) +
  theme_classic(base_size = 12) +
  theme(
    axis.title = element_text(face = "bold"), 
    axis.text = element_text(color = "black", size = rel(0.9)),
    axis.text.x = element_text(angle = 45, hjust = 1), 
    axis.ticks.x = element_blank()
  )
  
```


> G. What kind of graph did you create? Why?

I chose to use a bar graph because bar graphs can be used to show the distribution of one categorical variable, in this case, bird order

## Links to peer review

I reviewed Tyler Edvall's lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-tdedvall

Tyler Edvall reviewed my lab report at this URL: https://github.com/BIOL275-MSUM/lab-3-intro-to-ggplot-kelseyaleach/issues

## Session Info

Here is my session information:

```{r}
sessioninfo::session_info()
```
